---
title: "An Illustrative Data Analysis"
author: "Carlos Paniagua"
date: "9/8/2021"
output: html_document
---

```{r,include=FALSE}
library(tidyverse)

```

## Types of Movies Your Favorite Actor Makes

Inspired by the [Hollywood Taxonomy by Walt Hickey at fivethirtyeight.com](https://fivethirtyeight.com/tag/hollywood-taxonomy/).

### Idris Elba

-   We will do a similar analysis for [Idris Elba](https://www.rottentomatoes.com/celebrity/idris_elba)

-   We will use a clustering algorith to classify the different types of movies this actor makes

-   Goal: Write an application that perform similarly for any other actor

### Step 1: Get the data!

-   We will use movie ratings data from [Rotten Tomatoes](https://www.rottentomatoes.com/celebrity/idris_elba).

-   Getting data from a webpage is called ***scrapping**.*

#### Let's get the ratings first!

```{r, echo=FALSE}
idrisMovies = read.csv('idris_movies.csv')
idrisMovies %>% select(c(1,3,6)) %>% head()
```

The original dataset includes US domestic gross information but we will get this from another source.

#### Let's get the movie budgets and revenue!

We will scrape data from [the-numbers.com](https://www.the-numbers.com/movie/budgets/all/101).

```{r, echo=FALSE}
movies = read.csv('movies.csv')
head(movies)

```

Question: Are the `DomesticGross` and `WorldwideGross` columns adjusted for inflation?

```{r, echo=FALSE}
df = movies
df = df[complete.cases(df),]
year = str_split_fixed(df$ReleaseDate,", ",2)[,2]
df$year = year
df = df %>% relocate(year,Movie,ProductionBudget,DomesticGross,WorldwideGross)

dollar_to_int = function(x){
    gsub('[\\$,]',"", x) %>% as.integer()
}

percent_to_int = function(v){
    as.integer(sub("%",'',v))
    }

movies_df = df %>%
    mutate(ProductionBudget = dollar_to_int(ProductionBudget),
           DomesticGross = dollar_to_int(DomesticGross),
           WorldwideGross = dollar_to_int(WorldwideGross))

# Question: Is DomesticGross adjusted for inflation?

movies_df %>% ggplot(aes(x=year,y=DomesticGross/1000000))+geom_boxplot()+
    theme(axis.text.x = element_text(size=6,angle = 90,vjust = 0.5, hjust=1))+
    ylab('Domestic gross (Millions)')

movies_df = rename(movies_df, Title = Movie)

```

Probably not. Can you see why?

## 

```{r cars}
summary(cars)
```

### Step 2: Data Wrangling

We have to combine these two datasets into one so we can analyse it. This is called *Data Wrangling* or *Data Munging*.

```{r, echo=FALSE}
idris_df = idrisMovies
merged = left_join(idris_df,movies_df,by='Title')
merged = merged[complete.cases(merged),]
idris = merged %>% select(c(1,3,4,5,6,8,9,10))
# print(names(idris))
idris = rename(idris, Rating = `TomatometerÂ.`,BoxOffice = `Box.Office`)
# idris = idris %>% mutate(Rating = percent_to_int(Rating))
idris = na.omit(idris)
idris %>% select(c(1,2,6)) %>% as_tibble()
# idris = idris %>% mutate(Rating = ) 

```

### Step 3: Visualize the data

Let us plot our data!

```{r, echo=FALSE}
idris = idris %>% mutate(Rating = percent_to_int(Rating))
p = idris %>% ggplot(aes(x=Rating,y=DomesticGross/1000000))+
    geom_point() +
    xlab('Rating') + 
    ylab('Domestic gross (Millions)')
p
```

### Step 4: Modeling the data

We will use a [clustering algorithm](https://en.wikipedia.org/wiki/Cluster_analysis) called [*k*-means clustering](https://en.wikipedia.org/wiki/K-means_clustering) to group Idris Elba's movies. To do this we must choose the number of groups (clusters). Five clusters seems a good choice. Here are the results:

```{r, echo=FALSE}
actor = 'Idris Elba'
dfc = idris %>% select(Rating,DomesticGross)
k5 = kmeans(dfc,centers=5, nstart =  25)
cluster = factor(k5$cluster)
p5=idris %>% ggplot(aes(x=Rating,y=DomesticGross/1000000,color=cluster))+
    geom_point() +
    xlab('Rating') + 
    ylab('Domestic gross (Millions)')+
    ggtitle(paste0(actor,' Movies'))+
    theme(plot.title = element_text(hjust = 0.5))
    
p5

```

Here is the finished dataset including the clusters.

```{r, echo=FALSE}
idris_short = idris %>% select(Title,Rating,DomesticGross) %>% mutate(cluster = k5$cluster)
idris_short

```
