---
title: "Visualizing data"
author: "Carlos Paniagua"
date: "10/24/2021"
footer: "DSA201 Fall 2021, Carlos Paniagua"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)

# Install the RSocrata, sqldf packages if not available

if (!'palmerpenguins' %in% installed.packages()) install.packages("palmerpenguins")

# Load relevant packages
library(psych)
library(sqldf)
library(datasets)
library(tidyverse)
library(lubridate)
library(RColorBrewer)
library(palmerpenguins)


```

## Why represent data visually?

-   The eye is very powerful
-   Provide a concrete representation of the data
-   Insight may be hidden deeper than what numerical analysis can extract

## Motivating example: Anscombe's quartet

Here is simple data set with 8 variables known as Anscombe's quartet

```{r}
anscombe

```

## Summary statistics for Anscombe's quartet

```{r}
psych::describe(anscombe)[,2:4]
```

## Visualizing Anscombe's quartet

We use R base `plot` function to produce *scatter* plots for each pair of variables `x` and `y`.

```{r}
myplot = function(x,y){
    plot(x,y,pch=16,cex=1.5)
}


par(mfrow = c(1,4),pty="s")
with(anscombe,{
     myplot(x1,y1);
     myplot(x2,y2);
     myplot(x3,y3);
     myplot(x4,y4)}
     )






```

## Other visualizations using base R

Let us go back to the MD crashes data

```{r}

crashes_df <-read.csv('mdCrashes.csv')
crashes_df$Gdates = ymd(crashes_df$acc_date)
crashes_df$day_of_week = wday(crashes_df$Gdates)

crashes_per_day_week <- sqldf("
      
      select day_of_week, round(count(day_of_week) *100.0 / 
                          (select count(day_of_week) from crashes_df),2) as proportion
      from crashes_df
      group by day_of_week 
      ORDER BY proportion DESC;
      
      
      
      ")
crashes_per_day_week




```

## Pareto chart

We will produce a barplot with columns sorted from largest to smallest (Pareto chart)

```{r}
barplot(crashes_per_day_week$proportion, 
        main= 'Crashes in MD per day of week',
        xlab='day of week',
        ylab = 'proportion (%)',
        names.arg = crashes_per_day_week$day_of_week,
        col= brewer.pal(7, "Set2") )
```

## Box plots

-   Like histograms, box plots provide a visual for the distribtion of numerical data.
-   We will use the iris data set to look at the distribution `Petal.Length` for the three species available.

```{r}

data(iris)
head(iris)
```

## Box plots

```{r}
boxplot(Petal.Length ~ Species, data = iris,
        main='Petal length by species',
        col = brewer.pal(3, "Set3") )
```

## Histograms

```{r}

setosa_petal_length = iris$Petal.Length[iris$Species == 'setosa']
hist(setosa_petal_length)
```

## More advanced vizs with `ggplot2`

    library(ggplot2)

    # Standard recipe

    ggplot(data = [dataset], 
           mapping = aes(x = [x-variable], 
                         y = [y-variable]),
                         [other options]) +
       geom_xxx() +
       other options

## Using `ggplot2` to visualize `penguins` {.tabset .tabset-fade .tabset-pills}

### The `penguins` data set

```{r, include=TRUE, echo=TRUE}
glimpse(penguins)

```

### Code

    ggplot(data = penguins, 
           mapping = aes(x = bill_depth_mm, y = bill_length_mm,
                         colour = species)) +
      geom_point() +
      labs(title = "Bill depth and length",
           subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
           x = "Bill depth (mm)", y = "Bill length (mm)",
           colour = "Species")

::: {style="text-align: right"}
[[See Visualizing data with ggplot2](Visualising%20data%20with%20ggplot2)]
:::

## Viz

```{r}

# Start with the penguins data frame, map bill depth to the x-axis and map bill length to the y-axis.
ggplot(data = penguins, 
       mapping = aes(x = bill_depth_mm,
                     y = bill_length_mm,
                     # map species to the color of each point  
                     color = species)) + 
    
# Represent observation with a point    
  geom_point() +
    
   
  labs(
      
      # title data goes here
      title = "Bill depth and length",
      
      # subtitle goes here
      subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
       
      # label for x axis
      x = "Bill depth (mm)",
      
      # and for the y axis
      y = "Bill length (mm)",
      
      # label for the legend
      colour = "Species",
      
      # add a caption
      caption =  "Source: Palmer Station LTER / palmerpenguins package"
      ) +
    
    # use discrete color scale for accessibility
    scale_color_viridis_d()
```

## Aesthetics Options

Commonly used characteristics of chart attributes that can be mapped to a specific variable in the data are

-   color

-   shape

-   size

-   alpha (transparency)

## Shape

```{r, warning=FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm, 
           y = bill_length_mm,
           color = species,
           shape = island)) +
  geom_point() +
  scale_colour_viridis_d()
```

```{r, warning=FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm, 
           y = bill_length_mm,
           color = species,
           shape = species)) +  #same variable as color
  geom_point() +
  scale_colour_viridis_d()
```

## Size

```{r, warning=FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm, 
           y = bill_length_mm,
           color = species,
           shape = species,
           size = body_mass_g)) +
  geom_point() +
  scale_colour_viridis_d()
```

## Alpha (transparency)

```{r, warning=F}
ggplot(penguins,
       aes(x = bill_depth_mm, 
           y = bill_length_mm,
           colour = species,
           shape = species,
           size = body_mass_g,
           alpha = flipper_length_mm)) +
  geom_point() 

```
